---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# enaR: Tools for Ecological Network Analysis

<!-- badges: start -->
<!-- badges: end -->

[![DOI](https://zenodo.org/badge/12623293.svg)](https://zenodo.org/badge/latestdoi/12623293)

## Installation

You can install the development version of enaR like so:

``` r
remotes::install_github("SEELab/enaR")
```

## Basic Usage

Load a model from our set of models included with the package and
generate the full set of ENA metrics and indicators:

```{r example}
library(enaR)
data(enaModels)
model <- enaModels[[8]]
model.ena <- enaAll(model)
```

You can now explore the many metrics produced for the model, such as
structural aspects of the model:

```{r metrics}
model.ena$structure
```

For a more in-depth introduction to ENA and how to use the
[enaR](https://cran.r-project.org/package=enaR) package, you can view
our [website](url)