---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# enaR: Tools for Ecological Network Analysis

<!-- badges: start -->
[![pkgdown](https://github.com/SEELab/enaR/actions/workflows/pkgdown.yaml/badge.svg)](https://github.com/SEELab/enaR/actions/workflows/pkgdown.yaml)
<!-- badges: end -->

[![DOI](https://zenodo.org/badge/12623293.svg)](https://zenodo.org/badge/latestdoi/12623293)


`enaR` provides a set of high-level functions for conducting ecological network analysis (ENA). These functions allow users to access the many tools developed over decades of work by ecologists looking for ways to measure aspects of the struture and functioning of complex ecological systems, such as food-webs or biogeochemical cycles. 

In addition to collecting the many ENA algorithms together into a
single, open-source toolbox,`enaR` also facilitates the import, construction and simulation of ecological network models. There are multiple functions for reading in data from many of the various model formats that have arisen over the years.

## Installation

You can install the current release `enaR` like so:

``` r
remotes::install_github("SEELab/enaR")
```

## Basic Usage

Load a model from our set of models included with the package and
generate the full set of ENA metrics and indicators:

```{r example}
library(enaR)
data(enaModels)
model <- enaModels[[8]]
model.ena <- enaAll(model)
```

You can now explore the many metrics produced for the model, such as
structural aspects of the model:

```{r metrics}
model.ena$structure
```

For a more in-depth introduction to ENA and how to use the
[enaR](https://cran.r-project.org/package=enaR) package, you can view
our [website](https://seelab.github.io/enaR/)
