---
title: "Entering your Data"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Entering your Data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(enaR)
```

# Model

`enaR` is a software suite to apply Ecological Network Analyses to models that trace the flow of energy or matter through ecosystems.  The software does not currently include functions to help investigators construct a model; it assumes that you have a model in hand to analyze.  Investigators might build a model using a phenomenological approach or using systems dynamics modeling. It is also possible to use software tools to aid in this model construction like EcoPath with EcoSim (EwE) or limSolve for Linear Inverse Modeling.  Regardless of the model construction technique, the full set of information required for ENA includes the following:

* flow from node i to node j, which is typically summarized in a flow matrix **F**.  In a trophic ecosystem model, many of these flows will be generated by feeding relationships.  The Diet matrix in EcoPath contains some of this information.
* boundary input flows from the environment to node i, summarized as an input vector **z**.
* boundary losses from node i.  This can be summarized as a single vector (**y**), but in some cases it is necessary to distinguish between two types of boundary losses:
 * respirations losses refer to low quality losses that cannot be used to do further work (**r**).  
 * export losses refer to lost energy or material that can be used to do further work in other systems (**e**).   
* Storage (e.g., biomass) for each node (**X**)
* Living is a logical vector that indicates whether not a node is a living entity (TRUE) or not (FALSE).  This critical for some of the analysis algorithms (e.g., Mixed Trophic Impacts _enaMTI()_) that treat living and non-living nodes differently.  
* Node Names is a vector of the names of the nodes.  While this is not strictly necessary, it makes reading the enaR output much easier. 

In summary, a fully specified model for enaR includes {**F**, **z**, **r**, **e**, **y** = **r** + **e**, **X**, **Living**, **Names**}.  To be clear, however, not all analyses require all of this information. 

# Entering Data

The _pack()_ function can be used to create the `enaR` required data object.  This assumes the you have the necessary model data and that you can enter it or get it directly into R.   

# Reading Other Data Formats

Many investigators have their ecological network models in one of several formats.  The `enaR` package provides several functions that read in the data from these formats that then construct the `enaR` network data object.  Existing read functions include:

* _read.scor()_ reads in a plain text file of data formatted using the SCOR format based on sparse matrices that was developed as the input for Dr. Ulanowicz's NETWRK software.
* _read.nea()_ that reads in the NEA.m formatted system matrix stored as a comma separated file.
* _read.wand()_ reads the data from an MS Excel spreadsheet formatted for the WAND program.  
* _read.EcoNet()_ constructs the `enaR` network data object from data formatted as input for the EcoNet program.
* _read.enam()_ reads in the data from an extended matrix format used by some colleagues.  

As the most commonly used model construction software for food webs is EcoPath (EwE), it would be ideal to have a function to read in data from this format.  This is technically more challenging, but a goal for the future.    


 