---
title: "Flow Analysis: Simple"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Flow Analysis: Simple}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  warning = F,
  message = F
)
```

```{r setup}
library(enaR)
library(network)
```


The `enaR` library can be used to perform the Flow Analysis of Ecological Network Analysis. This includes finding the node throughflows, calculating the Input and Output analyses, and determining a set of whole-network metrics.

## Load the data

```{r flowsim1, eval=T}
# load a model
data(enaModels)
m <- enaModels[[9]] # select the oyster reef model
```

## Perform the analysis

```{r flowsim2, eval=T}
f <- enaFlow(m)  # peform the ENA flow analysis
attributes(f)
show(f$N)  # integral output-oriented flow intensity

show(f$ns) # vector of flow-based network statisics

```

##Ascendency Metrics

The ascendancy metrics proposed by Dr. Ulanowicz are also most often applied to the network flow distributions. In enaR this is done as follows.

```{r flowsim3, eval=T}
a <- enaAscendency(m)  # calculate the Ascendnecy metrics
show(a)
```