---
title: "Vizualizing your Network"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vizualizing your Network}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  warning = F,
  message = F
)
```

```{r setup}
library(enaR)
library(network)
```

This page illustrates how to use visualize a network model using the tools in _enaR_ and the _network_ package. 


Load the library of models and select one to use for this illustration.

```{r vizual1, eval = T}
# load data
data(enaModels)        # load library of Ecosystem Networks
names(enaModels)       # view model names
m <- enaModels[[9]]    # select the oyster model
```

## Simple Plot

We can then create a simple plot of the model with the default graphics parameters as

```{r vizual2, eval = T}
# Set the random seed to control plot output
set.seed(2)  # this is only used to ensure our plots look the same. 

# Plot network data object (uses plot.network)
plot(m)
```

## Fancy Plot

We can create a fancier plot by exerting more control on the plotting parameters.  Here is an example.

```{r vizual3, eval = T}

  ## Set colors to use
  my.col <- c('red','yellow',rgb(204,204,153,maxColorValue=255),'grey22')
  ## Extract flow information for later use.
  F <- as.matrix(m,attrname='flow')
  ## Get indices of positive flows
  f <- which(F!=0, arr.ind=T)
  opar <- par(las=1,bg=my.col[4],xpd=TRUE,mai=c(1.02, 0.62, 0.82, 0.42))
  ## Set the random seed to control plot output
  set.seed(2)
  plot(m,
  ## Scale nodes with storage
       vertex.cex=log(m%v%'storage'),
  ## Add node labels
       label= m%v%'vertex.names',
       boxed.labels=FALSE,
       label.cex=0.65,
  ## Make rounded nodes
       vertex.sides=45,
  ## Scale arrows to flow magnitude
       edge.lwd=log10(abs(F[f])),
       edge.col=my.col[3],
       vertex.col=my.col[1],
       label.col='white',
       vertex.border = my.col[3],
       vertex.lty = 1,
       xlim=c(-4,1),ylim=c(-2,-2))
  
```